Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-02-06 vesicle_code/docs/adhesion/adhesion.tex
removed todo comments

\section{Introduction}
Vesicles (closed fluid-filled phospholipid bilayer membranes) have been
widely utilized as biological cell mimics in biophysics and material
engineering \cite{sackmann1996,Barthes-Biesel2016_ARFM}.  The ever expanding applications have
encouraged detailed experimental, theoretical, and numerical
investigations of vesicle hydrodynamics in applied flow, electric
fields, and magnetic fields.  Theoretical investigations often take the
form of small-deformation analysis or spheroidal dynamics of a
nearly-spherical vesicle (or nearly-circular in two dimensions)
subjected to an applied flow or under an external
forcing~\cite{Barthes-BieselRallison1981_JFM, Misbah2006_PRL,
Vlahovska2007_PRE, Finken2008_EPL, ZhangZahnTanLin2013_PoF,
Nganguia2013_PRE}.  On the other hand, various numerical methods have
been developed for simulating the transient hydrodynamics of vesicle
suspensions~\cite{BagchiJohoson2005_JBE, Biben2005_EJP,
Veerapaneni2009_JCP, SeolHuKimLai2016_JCP}.

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by the reduced volume (reduced area in two dimensions),
viscosity contrast between interior and exterior fluids, and shear rate
of the imposed far-field fluid flow. In addition, a vesicle with a rigid
particle inside is also investigated as a biological mimic of a
eukaryotic cell with a nucleus that occupies nearly half of the
intracellular volume~\cite{Veerapaneni2011_PRL}. Small-deformation
analysis shows that a vesicle tank-treads in a planar shear flow for low
viscosity contrast and shear rate. At high viscosity contrast the
tank-treading dynamics transitions to tumbling 
dynamics~\cite{Misbah2006_PRL, Vlahovska2007_PRE}, and this leads to a
transition in effective shear viscosity in the vesicle
suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also validated
by direct numerical simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{DeschampsKantsler2009_PNAS,KantslerSegreSteinberg2008_EPL,
ZabuskySegreDeschamps2011_PoF}.  Between tank-treading and tumbling
vesicle dynamics, a breathing (tremble) mode is also
observed~\cite{Misbah2006_PRL,KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to affect the effective
shear viscosity in a different way.  In an extensional flow (planar or
uniaxial), vesicle shape dynamics depends sensitively on the vesicle
reduced volume and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape
and oscillatory undulation of the vesicle membrane are two examples of
the complex vesicle hydrodynamics in an extensional flow.

Collective hydrodynamics of vesicles is dictated by the vesicle-vesicle
interactions.  In a quiescent flow, vesicle-vesicle adhesion leads to
the formation of vesicle doublets~\cite{Ziherl2007_PRL,
ZiherlSvetina2007_PNAS} or clusters
\cite{SvetinaZiherl2008_Bioelectrochemistry,
FlormannAouane2017_SciReports}.  As a model for red blood cell (RBC)
aggregates, an idealized model for adhesive vesicle-vesicle interactions
gives rise to vesicle configurations similar to those observed in
experiments of fibrinogen-induced RBC
aggregates~\cite{SvetinaZiherl2008_Bioelectrochemistry,
FlormannAouane2017_SciReports}.  Using the Lennard-Jones (L-J) potential
for point-point interaction between two vesicle membranes without the
molecular details of adhesive interactions between RBCs, Flormann {\em
et al.}~found the contact zone to buckle under strong
adhesion~\cite{FlormannAouane2017_SciReports}.  A direct consequence of
the buckling is the sigmoidal shape of the adhesion region, which is
also observed in RBC doublets and explained in a slightly different
model~\cite{ZiherlSvetina2007_PNAS}.  It remains unknown how adhesive
interactions between vesicles might lead to different vesicle
hydrodynamics that results in different rheological properties of a
vesicle suspension. Such studies will provide useful insight to the
rheology of blood flow and how to use nano solutes to control the
hypertension by adjusting the adhesive interactions between RBCs.

@@

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.4]{figs/Dec13a_time_scaling01.png}
\includegraphics[keepaspectratio=true,scale=0.4]{figs/Dec13a_time_scaling02.png}
\caption{\label{fig:qflow00} Scaling of the time it takes for two
  vesicles to reach equilibrium. {\em Left}: Scaling with respect to
  $\delta$ with $\mathcal{H}=0.1$.  {\em Right}: Scaling with respect to
  $\mathcal{H}$ with
  $\delta = 0.4$. Arrows indicate the time by when the equilibrium is
  reached within $1\%$.}
  \todo[inline]{Change the $H$ in the right figure to $\mathcal{H}$}
\end{figure}

@@

In this work we focus on weakly adhesive vesicles whose equilibrium
shapes in a quiescent flow are shown in the left plot of
Figure~\ref{fig:Dec18_vesicle_shape} for three reduced areas: $\Delta
A=0.95$, $0.75$, and $0.6$ with $\mathcal{H}=5$ and $\delta = 0.2$. 
\begin{figure}
  \includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_vesicle_shape_vs_rA_00.jpeg}
  \includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_vesicle_shape_vs_rA_01.jpeg}
  \caption{\label{fig:Dec18_vesicle_shape} Two vesicles of the same length, reduced area
  and bending rigidity ($\kappa=1$)
  with $\mathcal{H}=5$, and $\delta=0.2$.  {\em Left}:
  Equilibrium shapes of two vesicles under adhesive interactions with
  adhesion strength $\mathcal{H}=5$ and separation distance
  $\delta=0.2$.  Vesicle reduced area $\Delta A=0.95$, $0.75$, and $0.6$
  are labeled.  The color coding is the tension along the vesicle. {\em
  Right}: Membrane shapes in the contact region.} \todo[inline]{This
  caption contradicts itself.  Is it $\mathcal{H} = 0.5$ or $\mathcal{H}
  = 5$?}
\end{figure}
The equilibrium vesicle shape for $\Delta A=0.95$ is the circular cap
with a flat contact region.  This is similar to the observed shapes of
two vesicles under strong adhesive interaction
in~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.  When
vesicles are more deflated with a reduced area $\Delta A = 0.75$, the
equilibrium vesicle shape is elongated with a bigger contact region.
This is consistent with the equilibrium shapes of a vesicle doublet
under a L-J potential ($(m,n) =
(12,6)$)~\cite{FlormannAouane2017_SciReports}.  As the reduced area
decreases further, we observe undulation of the elongated equilibrium
vesicle shape on the non-contact side while the contact region remains
flat.  The color coding along each curve is the tension of the vesicle
membrane.  We observe that the membrane tension in the contact region is
very negative, indicating a dominant compression of membrane when the
adhesion force is strong to keep the vesicles bound together.

@@

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.18]{figs/Dec18_Eb_vs_sigma_rA0p9502.jpeg}
\includegraphics[keepaspectratio=true,scale=0.18]{figs/Dec18_Eadh_vs_sigma_rA0p9502.jpeg}
  \caption{\label{fig:Dec18_equilibrium} Total bending energy (\em left)
  and adhesion energy {\em right} versus the adhesion strength
  $\mathcal{H}$ for two vesicles with reduced area $\Delta A=0.95$,
  separation distance $\delta = 0.2$ (triangles), $0.3$ (crosses), and
  $0.4$ (circles).}
  \todo[inline]{Change $H$ to $\mathcal{H}$ in the x-axis.}
\end{figure}

@@

We also examine how adhesive interaction may lead to formation of a
vesicle doublet dynamically in flowing conditions.  We simulate two
vesicles approaching each other in a planar shear flow, and examine how
their adhesive interactions lead to doublet formation.  Once a doublet
forms, the two vesicle membranes rotate around each other as they deform
dynamically. The usual tank-treading motion of a vesicle under shear
flow is not observed in each of the two vesicles. We compute the
effective shear viscosity of a dilute suspension of vesicle doublets,
and found it to be more than twice the effective shear viscosity of a dilute suspension of single vesicles.
Furthermore, we find that the membrane adhesion contribution to the shear viscosity increases with decreasing
reduced area while the bending/tension contribution increases with the reduced area.
%
%because deflated vesicles (low reduced area) deform more
%easily and more membranes in the contact region.
\todo[inline]{I don't understand this last sentence --> modified now by yny}

>>>>

\section{Introduction}
Vesicles (closed fluid-filled phospholipid bilayer membranes) have been
widely utilized as biological cell mimics in biophysics and material
engineering~\cite{sackmann1996, Barthes-Biesel2016_ARFM}.  The ever
expanding applications have encouraged detailed experimental,
theoretical, and numerical investigations of vesicle hydrodynamics in
applied flow, electric fields, and magnetic fields.  Theoretical
investigations often take the form of small-deformation analysis or
spheroidal dynamics of a nearly-spherical vesicle (or nearly-circular in
two dimensions) subjected to an applied flow or under an external
forcing~\cite{Barthes-BieselRallison1981_JFM, Misbah2006_PRL,
Vlahovska2007_PRE, Finken2008_EPL, ZhangZahnTanLin2013_PoF,
Nganguia2013_PRE}.  On the other hand, various numerical methods have
been developed for simulating the transient hydrodynamics of vesicle
suspensions~\cite{BagchiJohoson2005_JBE, Biben2005_EJP,
Veerapaneni2009_JCP, SeolHuKimLai2016_JCP}.

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by the reduced volume (reduced area in two dimensions),
viscosity contrast between interior and exterior fluids, and shear rate
of the imposed far-field fluid flow. In addition, a vesicle with a rigid
particle inside is also investigated as a biological mimic of a
eukaryotic cell with a nucleus that occupies nearly half of the
intracellular volume~\cite{Veerapaneni2011_PRL}. Small-deformation
analysis shows that a vesicle tank-treads in a planar shear flow for low
viscosity contrast and shear rate. At high viscosity contrast the
tank-treading dynamics transitions to tumbling
dynamics~\cite{Misbah2006_PRL, Vlahovska2007_PRE}, and this leads to a
transition in effective shear viscosity in the vesicle
suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also validated
by direct numerical simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{DeschampsKantsler2009_PNAS,
KantslerSegreSteinberg2008_EPL, ZabuskySegreDeschamps2011_PoF}.  Between
tank-treading and tumbling vesicle dynamics, a breathing (tremble) mode
is also observed~\cite{Misbah2006_PRL, KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to affect the effective
shear viscosity in a different way.  In an extensional flow (planar or
uniaxial), vesicle shape dynamics depends sensitively on the vesicle
reduced volume and the elastic capillary
number~\cite{KantslerSegreSteinberg2008_PRL, ZhaoShaqfeh2013_JFM,
Narsimhan2014_JFM, DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape
and oscillatory undulation of the vesicle membrane are two examples of
the complex vesicle hydrodynamics in an extensional flow.

Collective hydrodynamics of vesicles is dictated by the vesicle-vesicle
interactions.  In a quiescent flow, vesicle-vesicle adhesion leads to
the formation of vesicle doublets~\cite{Ziherl2007_PRL,
ZiherlSvetina2007_PNAS} or
clusters~\cite{SvetinaZiherl2008_Bioelectrochemistry,
FlormannAouane2017_SciReports}.  As a model for red blood cell (RBC)
aggregates, an idealized model for adhesive vesicle-vesicle interactions
gives rise to vesicle configurations similar to those observed in
experiments of fibrinogen-induced RBC
aggregates~\cite{SvetinaZiherl2008_Bioelectrochemistry,
FlormannAouane2017_SciReports}.  Using the Lennard-Jones (L-J) potential
for point-point interaction between two vesicle membranes without the
molecular details of adhesive interactions between RBCs, Flormann {\em
et al.}~found the contact zone to buckle under strong
adhesion~\cite{FlormannAouane2017_SciReports}.  A direct consequence of
the buckling is the sigmoidal shape of the adhesion region, which is
also observed in RBC doublets and explained in a slightly different
model~\cite{ZiherlSvetina2007_PNAS}.  It remains unknown how adhesive
interactions between vesicles might lead to different vesicle
hydrodynamics that results in different rheological properties of a
vesicle suspension. Such studies will provide useful insight to the
rheology of blood flow and how to use nano solutes to control the
hypertension by adjusting the adhesive interactions between RBCs.

@@

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.4]{figs/Dec13a_time_scaling01.png}
\includegraphics[keepaspectratio=true,scale=0.4]{figs/Dec13a_time_scaling02.png}
\caption{\label{fig:qflow00} Scaling of the time it takes for two
  vesicles to reach equilibrium. {\em Left}: Scaling with respect to
  $\delta$ with $\mathcal{H}=0.1$.  {\em Right}: Scaling with respect to
  $\mathcal{H}$ with
  $\delta = 0.4$. Arrows indicate the time by when the equilibrium is
  reached within $1\%$.}
\end{figure}

@@

In this work we focus on weakly adhesive vesicles whose equilibrium
shapes in a quiescent flow are shown in the left plot of
Figure~\ref{fig:Dec18_vesicle_shape} for three reduced areas: $\Delta
A=0.95$, $0.75$, and $0.6$ with $\mathcal{H}=5$ and $\delta = 0.2$. 
\begin{figure}
  \includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_vesicle_shape_vs_rA_00.jpeg}
  \includegraphics[keepaspectratio=true,scale=0.175]{figs/Dec18_vesicle_shape_vs_rA_01.jpeg}
  \caption{\label{fig:Dec18_vesicle_shape} Two vesicles of the same length, reduced area
  and bending rigidity ($\kappa=1$)
  with $\mathcal{H}=5$, and $\delta=0.2$.  {\em Left}:
  Equilibrium shapes of two vesicles under adhesive interactions with
  adhesion strength $\mathcal{H}=5$ and separation distance
  $\delta=0.2$.  Vesicle reduced area $\Delta A=0.95$, $0.75$, and $0.6$
  are labeled.  The color coding is the tension along the vesicle. {\em
  Right}: Membrane shapes in the contact region.}
\end{figure}
The equilibrium vesicle shape for $\Delta A=0.95$ is the circular cap
with a flat contact region.  This is similar to the observed shapes of
two vesicles under strong adhesive interaction
in~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}.  When
vesicles are more deflated with a reduced area $\Delta A = 0.75$, the
equilibrium vesicle shape is elongated with a bigger contact region.
This is consistent with the equilibrium shapes of a vesicle doublet
under a L-J potential ($(m,n) =
(12,6)$)~\cite{FlormannAouane2017_SciReports}.  As the reduced area
decreases further, we observe undulation of the elongated equilibrium
vesicle shape on the non-contact side while the contact region remains
flat.  The color coding along each curve is the tension of the vesicle
membrane.  We observe that the membrane tension in the contact region is
very negative, indicating a dominant compression of membrane when the
adhesion force is strong to keep the vesicles bound together.

@@

\begin{figure}
\includegraphics[keepaspectratio=true,scale=0.18]{figs/Dec18_Eb_vs_sigma_rA0p9502.jpeg}
\includegraphics[keepaspectratio=true,scale=0.18]{figs/Dec18_Eadh_vs_sigma_rA0p9502.jpeg}
  \caption{\label{fig:Dec18_equilibrium} Total bending energy (\em left)
  and adhesion energy {\em right} versus the adhesion strength
  $\mathcal{H}$ for two vesicles with reduced area $\Delta A=0.95$,
  separation distance $\delta = 0.2$ (triangles), $0.3$ (crosses), and
  $0.4$ (circles).}
\end{figure}

@@

We also examine how adhesive interaction may lead to formation of a
vesicle doublet dynamically in flowing conditions.  We simulate two
vesicles approaching each other in a planar shear flow, and examine how
their adhesive interactions lead to doublet formation.  Once a doublet
forms, the two vesicle membranes rotate around each other as they deform
dynamically. The usual tank-treading motion of a vesicle under shear
flow is not observed in each of the two vesicles. We compute the
effective shear viscosity of a dilute suspension of vesicle doublets,
and found it to be more than twice the effective shear viscosity of a
dilute suspension of single vesicles.  Furthermore, we find that the
membrane adhesion contribution to the shear viscosity increases with
decreasing reduced area while the bending/tension contribution increases
with the reduced area.
